Kindly Bent to Free Us

### 2019-11-14 kindly/infer/extra.tex
Fixed uses of 'note' instead of 'denote' and various typos.
Add a macro for input parameters.

\subsection{Kind inference}
\label{rules:kinding:full}
%
\begin{figure}[tb]
  \centering
  \input{infer/kinds}
  \caption{Kind inference rules -- $\inferK{(C,\unif)}{{\bm\E}}{{\bm\tau}}{k}$}
  \label{rules:kinding}
\end{figure}
%
We note $\inferK{(C,\unif)}{{\bm\E}}{{\bm\tau}}{k}$ when type $\tau$ has kind $k$
in environment $\E$ under constraints $C$ and unifier $\unif$.
$\bm\E$ and $\bm\tau$ are the input parameters of
our inference procedure.
We present the kind inference algorithm as a set of rules in
\cref{rules:kinding}.
Higher-kinds are not generally supported
and can only appear by looking-up the kind scheme of a type constructor,
for use in the type application rule {\sc KApp}.
Type variables must be of a simple kind in rule {\sc KVar}.
Kind schemes are instantiated in the {\sc KVar} rules by creating
fresh kind variables and the associated substitution.
{\sc KArr} and {\sc KBorrow} simply returns the kind of the primitive
arrow and borrow types.
The $\operatorname{normalize}$ function is used every time several constraints
must be composed in order to simplify the constraint and return a most general
unifier.

\begin{figure*}[!btp]
  \input{infer/typing}
  \caption{Type inference rules --
    $\inferW{\Sigma}{(C,\psi)}{{\bm\E}}{{\bm e}}{\tau}$ }
  \label{rules:typing:full}
\end{figure*}

>>>>

\subsection{Kind inference}
\label{rules:kinding:full}
%
\begin{figure}[tb]
  \centering
  \input{infer/kinds}
  \caption{Kind inference rules -- $\inferK{(C,\unif)}{{\inP\E}}{{\inP\tau}}{k}$}
  \label{rules:kinding}
\end{figure}
%
We write $\inferK{(C,\unif)}{{\inP\E}}{{\inP\tau}}{k}$ when type $\tau$ has kind $k$
in environment $\E$ under constraints $C$ and unifier $\unif$.
$\inP\E$ and $\inP\tau$ are the input parameters of
our inference procedure.
We present the kind inference algorithm as a set of rules in
\cref{rules:kinding}.
Higher-kinds are not generally supported
and can only appear by looking-up the kind scheme of a type constructor,
for use in the type application rule {\sc KApp}.
Type variables must be of a simple kind in rule {\sc KVar}.
Kind schemes are instantiated in the {\sc KVar} rules by creating
fresh kind variables and the associated substitution.
{\sc KArr} and {\sc KBorrow} simply returns the kind of the primitive
arrow and borrow types.
The $\operatorname{normalize}$ function is used every time several constraints
must be composed in order to simplify the constraint and return a most general
unifier.

\begin{figure*}[!btp]
  \input{infer/typing}
  \caption{Type inference rules --
    $\inferW{\Sigma}{(C,\psi)}{{\inP\E}}{{\inP e}}{\tau}$ }
  \label{rules:typing:full}
\end{figure*}

### 2019-11-14 kindly/infer/infer.tex
Add a macro for input parameters.

In \cref{sdtyping}, we used relations to split environments in two and to
transform suspended bindings into borrows inside a region.
These relations take as argument a constraint which validates
the transformations.
In the context of inference, we define new judgements which \emph{infer}
the constraints.
\begin{itemize}[leftmargin=*]
\item $\bsplit{C}{\Sv}{\bm{\Sv_1}}{\bm{\Sv_2}}$.
  Given two usage environments $\Sv_1$ and $\Sv_2$
  (inferred from two subexpressions, for instance),
  we return $\Sv$, the merged environment, and $C$, a set
  of constraints that must be respected.
  This relation is total and non-ambiguous given $\Sv_1$ and $\Sv_2$
  and uses the same rules as the one presented in \cref{sdtyping:split}.
\item $\bregion[\bm{n}]{C}{\bm{x}}{\Sv}{\bm\Sv'}$.
  Given a usage environment $\Sv'$ and a variable name $x$, we return
  $\Sv$ where the borrow binding of $x$ in $\Sv'$, if it exists, is replaced by
  a suspended binding. We also return $C$, a set of constraints that must
  be respected.
  Again, the relation is total and non-ambiguous for any given $\Sv'$ and $x$,
  and uses the rules presented in \cref{sdtyping:regions}.
\end{itemize}

@@

We write $\inferW{\Sv}{(C,\unif)}{\bm{\E}}{\bm{e}}{\tau}$ when
$e$\ has type $\tau$ in $\E$ under the constraints $C$ and unifier $\unif$
with a usage environment $\Sv$. $\bm\E$ and $\bm e$ are the input parameters of our
inference algorithm.
Unlike in the syntax-directed version, $\E$ only regular and type bindings.
Suspended and borrow bindings can only be present in $\Sv$.
We revisit some of the syntax-directed rules presented in \cref{sdtyping}
to highlight the novelties
of our inference algorithm and the differences with the syntax-directed
system.
As before, the complete type inference rules are shown in \cref{appendix:infer}.

@@

    \ruleIRegion
  \end{mathpar}
  \caption{The {\sc Pair$_I$} and {\sc Region$_I$} rules}
  \label{rule:infer:envrules}
\end{figure}
%
We now present the inference version
of the {\sc Pair} and {\sc Region} rules in \cref{rule:infer:envrules}.
While the rules are quite similar to the original version, the usage
environment $\Sv$ is now an \emph{output} of the algorithm.
As such, we use the ``inference'' version of the relations on
the environment,
$\bsplit{C}{\Sv}{\bf{\Sv_1}}{\bf{\Sv_2}}$ and $\bregion{C}{\bf{x}}{\Sv}{\bf\Sv'}$,
which returns the necessary constraints.
As appropriate for inference, these relations
are total and non ambiguous for their input parameters ($\Sv_1, \Sv_2$ and $\Sv'$, respectively).
We then collect all constraints and normalize them.

@@

\begin{figure*}[!bt]
  \centering
  \begin{minipage}{0.32\linewidth}
    \begin{align*}
      C &::= \Cleq{\tau_1}{\tau_2} \\
        &\mid \Cleq{k_1}{k_2} \tag{Inequalities} \\
        &\mid C_1 \Cand C_2  \tag{Conjunction}\\
        &\mid \Cproj{\tvar}{C}
        \mid \Cproj{\kvar}{C} \tag{Projection}
    \end{align*}
    \caption{The constraint language}
    \label{grammar:constraint}
  \end{minipage}\hfill
  \begin{minipage}{0.66\linewidth}
    \begin{mathpar}
      \inferrule
      {\forall i,\ \entail{C}{\Cleq{l_i}{k}}}
      {\entail{C}{\Cleq{\wedge_i l_i}{k}}}
      \and
      \inferrule
      {\forall i,\ \entail{C}{\Cleq{k}{l_i}}}
      {\entail{C}{\Cleq{k}{\vee_i l_i}}}
      \and
      \inferrule
      { \forall i,\ \entail{C}{\Ceq{\tau_i}{\tau_i}}\\
      }
      { \entail{C}{\Cleq{\tapp{t}{(\tau_i)}}{\tapp{t}{(\tau'_i)}}} }
      \and
      \inferrule{l \lk_\Lat l'}{\entail{}{\Cleq{l}{l'}}}
      \and
      \inferrule
      { \entail{C}{\Cleq{\tau'_1}{\tau_1}}\\
        \entail{C}{\Cleq{\tau_2}{\tau'_2}}\\
        \entail{C}{\Cleq{k}{k'}}
      }
      { \entail{C}{\Cleq{\tau_1\tarr{k}\tau_2}{\tau'_1\tarr{k'}\tau'_2}} }
    \end{mathpar}
    \caption{Base entailment rules -- $\entail{C}{D}$ }
    \label{rules:entail}
  \end{minipage}
\end{figure*}

>>>>

In \cref{sdtyping}, we used relations to split environments in two and to
transform suspended bindings into borrows inside a region.
These relations take as argument a constraint which validates
the transformations.
In the context of inference, we define new judgements which \emph{infer}
the constraints.
\begin{itemize}[leftmargin=*]
\item $\bsplit{C}{\Sv}{\inP{\Sv_1}}{\inP{\Sv_2}}$.
  Given two usage environments $\Sv_1$ and $\Sv_2$
  (inferred from two subexpressions, for instance),
  we return $\Sv$, the merged environment, and $C$, a set
  of constraints that must be respected.
  This relation is total and non-ambiguous given $\Sv_1$ and $\Sv_2$
  and uses the same rules as the one presented in \cref{sdtyping:split}.
\item $\bregion[\inP{n}]{C}{\inP{x}}{\Sv}{\inP\Sv'}$.
  Given a usage environment $\Sv'$ and a variable name $x$, we return
  $\Sv$ where the borrow binding of $x$ in $\Sv'$, if it exists, is replaced by
  a suspended binding. We also return $C$, a set of constraints that must
  be respected.
  Again, the relation is total and non-ambiguous for any given $\Sv'$ and $x$,
  and uses the rules presented in \cref{sdtyping:regions}.
\end{itemize}

@@

We write $\inferW{\Sv}{(C,\unif)}{\inP{\E}}{\inP{e}}{\tau}$ when
$e$\ has type $\tau$ in $\E$ under the constraints $C$ and unifier $\unif$
with a usage environment $\Sv$. $\inP\E$ and $\inP{e}$ are the input parameters of our
inference algorithm.
Unlike in the syntax-directed version, $\E$ only regular and type bindings.
Suspended and borrow bindings can only be present in $\Sv$.
We revisit some of the syntax-directed rules presented in \cref{sdtyping}
to highlight the novelties
of our inference algorithm and the differences with the syntax-directed
system.
As before, the complete type inference rules are shown in \cref{appendix:infer}.

@@

    \ruleIRegion
  \end{mathpar}
  \caption{The {\sc Pair$_I$} and {\sc Region$_I$} rules}
  \label{rule:infer:envrules}
\end{figure}
%
We now present the inference version
of the {\sc Pair} and {\sc Region} rules in \cref{rule:infer:envrules}.
While the rules are quite similar to the original version, the usage
environment $\Sv$ is now an \emph{output} of the algorithm.
As such, we use the ``inference'' version of the relations on
the environment,
$\bsplit{C}{\Sv}{\inP{\Sv_1}}{\inP{\Sv_2}}$
and $\bregion{C}{\inP{x}}{\Sv}{\inP\Sv'}$,
which returns the necessary constraints.
As appropriate for inference, these relations
are total and non ambiguous for their input parameters ($\Sv_1, \Sv_2$ and $\Sv'$, respectively).
We then collect all constraints and normalize them.

@@

\begin{figure*}[!bt]
  \centering
  \begin{minipage}{0.32\linewidth}
    \begin{align*}
      C &::= \Cleq{\tau_1}{\tau_2} \\
        &\mid \Cleq{k_1}{k_2} \tag{Inequalities} \\
        &\mid C_1 \Cand C_2  \tag{Conjunction}\\
        &\mid \Cproj{\tvar}{C}
        \mid \Cproj{\kvar}{C} \tag{Projection}
    \end{align*}
    \caption{The constraint language}
    \label{grammar:constraint}
  \end{minipage}\hfill
  \begin{minipage}{0.66\linewidth}
    \begin{mathpar}
      \inferrule
      {\forall i,\ \entail{C}{\Cleq{l_i}{k}}}
      {\entail{C}{\Cleq{\wedge_i l_i}{k}}}
      \and
      \inferrule
      {\forall i,\ \entail{C}{\Cleq{k}{l_i}}}
      {\entail{C}{\Cleq{k}{\vee_i l_i}}}
      \and
      \inferrule
      { \forall i,\ \entail{C}{\Ceq{\tau_i}{\tau_i}}\\
      }
      { \entail{C}{\Cleq{\tapp{\tcon}{(\tau_i)}}{\tapp{\tcon}{(\tau'_i)}}} }
      \and
      \inferrule{l \lk_\Lat l'}{\entail{}{\Cleq{l}{l'}}}
      \and
      \inferrule
      { \entail{C}{\Cleq{\tau'_1}{\tau_1}}\\
        \entail{C}{\Cleq{\tau_2}{\tau'_2}}\\
        \entail{C}{\Cleq{k}{k'}}
      }
      { \entail{C}{\Cleq{\tau_1\tarr{k}\tau_2}{\tau'_1\tarr{k'}\tau'_2}} }
    \end{mathpar}
    \caption{Base entailment rules -- $\entail{C}{D}$ }
    \label{rules:entail}
  \end{minipage}
\end{figure*}
